---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load the FAS_litters data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```

## `select` function

This function lets us choose some columns and not others
The colon specifies from variable X to variable Y

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

The minus specifies a variable to remove


```{r}
select(litters_df, -litter_number)
```

Renaming columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Using select helpers. Helps select columns want to keep based on some criteria of variable name.

```{r}
select(litters_df, starts_with("gd"))
```

Sometimes helpful to know everything will return everything. e.g. want to look at litter_number as first variable, and then also return rest of vars.

```{r}
select(litters_df, litter_number, everything())
```

Relocate function does same thing, re-orders columns with variable specified to beginning. 

```{r}
relocate(litters_df, litter_number)
```

## `filter` function

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

Because these two functions do not include all observations in df, shows that some of gd0_weight have NA in this variable.

More logical operators
Two equal signs is test for equality, rather than one equal sign that changes the variable. 

```{r}
filter(litters_df, gd_of_birth == 20)
```

The ! around == logical operator 
```{r}
filter(litters_df, !(gd_of_birth == 20))
```

```{r}
filter(litters_df, gd_of_birth != 20)
```

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

How to select multiple values of a variable. Where group is either Con7 or Mod8
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

## `mutate`

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```

## `arrange`

Can occasionally be helpful to arrange dataset in a certain way if want to examine it
```{r}
arrange(litters_df, pups_born_alive)
```

```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```

## `%>%`

What if want to import dataset, then clean names, then select some variables, then mutate and then filter... all at once?

What we do NOT want to do
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```

What to do instead - PIPE OPERATOR

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```

Can be more specific about where the results of steps in pipe operator will enter in subsequent code step. 