---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load the FAS_litters data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```

## `select` function

This function lets us choose some columns and not others
The colon specifies from variable X to variable Y

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

The minus specifies a variable to remove


```{r}
select(litters_df, -litter_number)
```

Renaming columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Using select helpers. Helps select columns want to keep based on some criteria of variable name.

```{r}
select(litters_df, starts_with("gd"))
```

Sometimes helpful to know everything will return everything. e.g. want to look at litter_number as first variable, and then also return rest of vars.

```{r}
select(litters_df, litter_number, everything())
```

Relocate function does same thing, re-orders columns with variable specified to beginning. 

```{r}
relocate(litters_df, litter_number)
```


