---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in csv data

Read in the litters dataset.

```{r litters}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df
litters_df = janitor::clean_names(litters_df)
```

## Look at the data

Printing in the console.
```{r}
litters_df
```

Looking at head and tail of df.
```{r}
head(litters_df)
tail(litters_df)
```

Skim function to look globally at data.
```{r}
skimr::skim(litters_df)
```

## Options to read cvs

These options can be helpful if data csv has blank initial rows, or if columns not variable names.
Therefore, can skip rows.
```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
litters_df
```

This can be helpful if need to recognize multiple missing values.
```{r}
litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", 999))
litters_df
```

Check out `?read_csv()` for more options. 

## Read in other file formats

Read in an excel file.

```{r}
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
```

Can also look at `?read_excel` for help. Especially focus on sheet = null or range = null if want to read-in specific cells.

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

Read in a SAS file.

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with Base R

What about `read.csv` ...?

```{r}
litters_base = read.csv("./data/FAS_litters.csv")
litters_readr = read_csv("./data/FAS_litters.csv")
```

Printed difference makes read_csv better. Also, how characters get parsed and incomplete variables handling etc. 
