---
title: "Tidy Data"
output: github_document
---

```{r setup}
library(tidyverse)
```

## `pivot_longer`

Load the PULSE data

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Note that bdi_score measured at baseline, 1 months, 6 months, 12 months.
Instead, want to have visit variable with values (baseline, 1 mon, 6 mon, 12 mon) and bdi score for each

Wide format to long format.

```{r}
pulse_data_tidy = 
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )
```

Rewrite, combine, and extend (to add a mutate step to code chunk)
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id, visit) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
```

## `pivot_wider`

Make up data

```{r}
analysis_result =
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )
```

Use Pivot wider

```{r}
analysis_result %>% 
  pivot_wider(
    names_from = "time",
    values_from = "mean"
  )
```

## Binding rows

Problem of merging data tables in LOTR word dataset. male/female should be values in single variable, but spread across 3 tables. 

Use LotR data.

Step 1: import each table.

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")
```

Bind all the rows together (stack these 3 datasets)

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>% 
  relocate(movie) %>% 
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words"
  )
```

